Light inextensible strings under tension in the Schwarzschild geometry

### 2018-10-01 string_microarticle.tex
rephrasing

\begin{abstract}
%%%
Here I analyze the behaviour of light inextensible strings under
tension in the vicinity of a black hole.  Proper length (given by the
spatial component of the Schwarzshild metric) is minimized by the
Euler-Lagrange equations.  Light inextensible taut strings may assume
a range of distinct configurations with an arbitrarily high winding
number around the central singularity, and here I give a range of
numerical solutions and theoretical observations.
%%%%
\end{abstract}

@@

\begin{equation}\label{schwarzschild}
ds^2= -dt^2\left(1-1/r\right) +\frac{dr^2}{1-1/r} + r^2\left(d\theta^2 + \sin^2\theta d\phi^2\right)
\end{equation}

where units in which~$2GM=c=1$ are used, has been known for over a
century~\cite{schwarzschild1916} and makes a number of
counter-intuitive predictions.  Here I consider the spatial
component~$ds^2= \frac{dr^2}{1-1/r} + r^2d\phi^2$, confined to
equatorial points~($\phi=\pi/2$); we understand~$1<r<\infty$
throughout.  A string of constant~$r$ will, trivially, have proper
length~$2\pi r$.

@@

where dashes denote differentiation with respect to~$\phi$.  The
Euler-Lagrange equations for this system,
$\frac{d}{d\phi}\frac{\partial F}{\partial r'}-\frac{\partial
  F}{\partial r}=0$ give, after simplification, an expression for the
second derivative of~$r$:

\begin{equation}\label{rdashdash}
  r''\left(\phi\right) =
  (r-1) + \frac{(4r-3)\left(r'\right)^2}{2r\left(r-1\right)},\qquad r>1
\end{equation}

This differential equation does not appear to have a simple analytical
solution; I present some numerical results below.  Solutions for
which~$r'(\phi)>0$ whenever~$r>1$ correspond to strings that have only
one free end; the other is tangential to the event horizon, except in
the exceptional case of constant~$\phi$ when the string is radial
for~$r>1$.  It is straightforward to show that the radius of curvature
of a taut string becomes infinite, and the string has a point of
inflection, if~$\left(r'\right)^2=2r\left(r-1\right)$.

@@

Figure~\ref{closest_approach_non_self_intersecting} shows
nonselfintersecting strings indexed by distance of approach to the
singularity, which occurs at~$\phi=0$.  Strings with a sufficiently
close passage to the event horizon self-intersect; numerical methods
show that the critical value is~$\simeq 1.076109317$, with an
uncertainty of 1 in the last digit: if a string with two free ends
passes closer than this, it forms a closed loop.
Figure~\ref{closest_approach_non_self_intersecting} shows some
non-self-intersecting strings and
figure~\ref{closest_approach_self_intersecting} shows some strings
that self-intersect.  Figure~\ref{one_free_end} shows some strings
that pass through~$\left(2,0\right)$ and tangent to the event horizon.

@@

The analysis above was restricted to the case~$r\geq 1$, where it is
possible to hold objects steady.  Of course, once within the event
horizon, even a light inextensible string cannot hold itself steady
(at least in Schwarzschild coordinates) and one would have to account
for the dynamics of thes system.

In figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the right, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted; the situation is
steady, so no gravitational waves are produced; the string is
inextensible, so stores no energy.  Likewise,
Figure~\ref{one_free_end} would suggest that angular momentum is
transmitted to the black hole, but no mechanism is apparent.

@@

>>>>

\begin{abstract}
%%%
Here I analyze the behaviour of light inextensible strings under
tension in the vicinity of a nonrotating black hole.  Proper length
(given by the spatial component of the Schwarzshild metric) is
minimized by the Euler-Lagrange equations.  Light inextensible taut
strings may assume topologically distinct configurations, and here I
give some numerical solutions and theoretical observations.
%%%%
\end{abstract}

@@

\begin{equation}\label{schwarzschild}
ds^2= -dt^2\left(1-1/r\right) +\frac{dr^2}{1-1/r} + r^2\left(d\theta^2 + \sin^2\theta d\phi^2\right),
\end{equation}

\noindent where units in which~$2GM=c=1$ are used, has been known for
over a century~\cite{schwarzschild1916} and makes a number of
counter-intuitive predictions.  Here I consider the spatial
component~$ds^2= \frac{dr^2}{1-1/r} + r^2d\phi^2$, confined to
equatorial points~($\theta=\pi/2$); we understand~$1<r<\infty$
throughout.  A string of constant~$r$ will, trivially, have proper
length~$2\pi r$.

@@

\noindent where dashes denote differentiation with respect to~$\phi$.  The
Euler-Lagrange equations for this system,
$\frac{d}{d\phi}\frac{\partial F}{\partial r'}-\frac{\partial
  F}{\partial r}=0$ give, after simplification, an expression for the
second derivative of~$r$:

\begin{equation}\label{rdashdash}
  r''\left(\phi\right) =
  (r-1) + \frac{(4r-3)\left(r'\right)^2}{2r\left(r-1\right)},\qquad r>1.
\end{equation}

This differential equation does not appear to have a simple analytical
solution; I present some numerical results below.  It is
straightforward to show that the radius of curvature of a taut string
becomes infinite, and the string has a point of inflection,
if~$\left(r'\right)^2=2r\left(r-1\right)$.

@@

Figure~\ref{closest_approach_non_self_intersecting} shows non-self
intersecting strings indexed by distance of approach to the
singularity, which is taken to occur at~$\phi=0$.  Strings with a
sufficiently close passage to the event horizon self-intersect;
numerical methods show that the critical value is~$\simeq
1.076109317$, with an uncertainty of 1 in the last digit: if the
closest approach is less than this value, the string crosses itself
at~$\phi=\pi/2$.  Figure~\ref{closest_approach_non_self_intersecting}
shows some non-self-intersecting strings and
figure~\ref{closest_approach_self_intersecting} shows some strings
that self-intersect.  Figure~\ref{one_free_end} shows some strings
that pass through~$\left(2,0\right)$ and are tangent to the event horizon.

@@

The analysis above was restricted to the case~$r\geq 1$, where it is
possible to hold objects steady.  Of course, once within the event
horizon, even a light inextensible string cannot hold itself steady
(at least in Schwarzschild coordinates) and one would have to account
for the dynamics of the system.

In figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the right, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted: the situation is
steady, so no gravitational waves are produced; the string is
inextensible, so stores no energy.  Likewise,
Figure~\ref{one_free_end} would suggest that angular momentum is
transmitted to the black hole, but no mechanism is apparent.

@@

\section*{Supplementary Material}

Calculations are performed using the R programming language; source
code for the figures is available at

\url{https://github.com/RobinHankin/string.git}
